---
title: "gcp-queries"
author: Alexander Nielson
output: html_document
---

# Set up correct Environment.

there is a known bug with the bigrquery library where the a numeric string is getting converted to scientific notation due to some change by the google developers. So until the R guys get a permanent fix, just run this line to turn off the scientific notation in R. 

```{r}
options(scipen = 999)
```

## Libs
```{r}
library(bigrquery)
library(DBI)
library(tidyverse)
library(data.table)
```

# Authhenticate / Connection
```{r}
#type bq_auth() into the console.

# Or if you have a service account token (reccomended), then simply put in the key path.
bq_auth(path = "D:/ut-sao-transparency-prod-3988539aa4d9.json")

project_name <- 'ut-sao-transparency-prod'

con <- dbConnect(
  bigrquery::bigquery(),
  project = project_name,
  dataset = "transaction",
  billing = project_name
)
```

```{r}
dbListTables(con)
```


#Query the entity_lookup table

this is a small table so querying it will not incur significant costs, so use it first to make sure you understand the process before you query the bigger table.

The process for querying can be broken down into three steps:

Step 1: Write your query  and assign as a string
```{r}
q <- "
      SELECT id, name, govt_lvl
      FROM `transaction.entity_lookup`
     "
```

Step 2: project the query costs. 
```{r}
projected_query <- bq_project_query(project_name, q)
```

Step 3: Execute the query.
```{r}
entity_lookup <- bq_table_download(projected_query)
```

Alternatively,if you already know the costs of a query and wish to simplify the steps, then you can run a query in one line using the standard DBI package.

```{r}
entity_lookup_2<- dbGetQuery(con, q)
```

# Querying the Transaction Table

```{r}
q <- "
      SELECT fiscal_year, type, vendor_name, amount, fund1, fund2, program1, program2, function1, function2
      FROM `transaction.transaction`
      WHERE entity_name ='South Sanpete School District' and type='EX'
      "
```

```{r}
dt_ssd <- dbGetQuery(con, q)
```

Lets get the yearly total up to two fund levels
```{r}
dt_ssd %>%
  group_by(fiscal_year,fund1,fund2) %>% 
  summarize(net_amount  = sum(amount))
```




# Saving query results and objects

Since everytime you query something you are charged (if you just ran the query its results will be stored as a temp table for a bit, but within a few hours or new session the costs will be incurred again), I reccomend that you save query results for further maniuplation.
```{r}
dt_ssd %>% saveRDS(paste0("dt_layton_",Sys.Date(),".RDS"))
```
or you can save it as a csv.

```{r}
dt_ssd %>% fwrite(paste0("dt_layton_",Sys.Date(),".csv"))
```



