---
title: "gcp-queries"
author: Alexander Nielson
output: html_document
---

# Set up correct Environment.

there is a known bug with the bigrquery library where the a numeric string is getting converted to scientific notation due to some change by the google developers. So until the R guys get a permanent fix, just run this line to turn off the scientific notation in R. 

```{r}
options(scipen = 999)
```

## Libs
```{r}
library(bigrquery)
library(DBI)
library(tidyverse)
```

# Authenticate to Big Query
```{r}
#Option 1: type bq_auth() into the console and log in with your email.

# Option2: If you have a service account token (reccomended), then simply put in the key path like this:
bq_auth(path = "D:/gcp-keys/ut-sao-transparency-prod-5a8476ec5fea.json")

project_name <- 'ut-sao-transparency-prod'

con <- dbConnect(
  bigrquery::bigquery(),
  project = project_name,
  dataset = "transaction",
  billing = project_name
)
```

# How to Query

The process for querying can be broken down into three steps:

Step 1: Write your query and assign as a string
```{r}
q <- "SELECT DISTINCT entity_name
FROM transaction.transaction
ORDER BY entity_name"
```

Step 2: Project the query costs. 
```{r}
projected_query <- bq_project_query(project_name, q)
```

Step 3: Execute the query and download
```{r}
entity_lookup <- bq_table_download(projected_query)
```

Alternatively,if you already know the costs of a query and wish to simplify the steps, then you can run a query in one line using the standard DBI package.

```{r}
entity_lookup_2 <- dbGetQuery(con, q)
```

# Example 2: More Complex Query

Lets imagine I want all the expense transaction from a school district ("South Sanpete School District" was used since it is small)

I want to see what funds, programs, and function codes have the highest spending. I would consult the table schema, then write my query like this:
```{r}
q <- "
      SELECT  fiscal_year, 
              type, 
              vendor_name, 
              amount, 
              fund1, 
              fund2, 
              program1, 
              program2, 
              function1, 
              function2
      FROM `transaction.transaction`
      WHERE entity_name ='South Sanpete School District' 
        AND type='EX'
      "
```

next I execute the query
```{r}
dt_ssd <- dbGetQuery(con, q)
```

Lets now find the net expense for each year and by eacj fund1 and fund2 level
```{r}
dt_ssd %>%
  group_by(fiscal_year, fund1, fund2) %>% 
  summarize(net_amount  = sum(amount)) 
```


# Saving Large Queries: 

Since everytime you query something you are charged, I recommend saving large queries to your local file system.

Saving your dataframe/tibble as an RDS object using "saveRDS" is encouraged. Here is an example: 
```{r}
dt_ssd %>% saveRDS(paste0("dt_layton_",Sys.Date(),".RDS"))
```

then to read it back in you can run "readRDS"
```{r}
dt_ssd <- readRDS(paste0("dt_layton_",Sys.Date(),".RDS"))
```


or you can save it as a csv.
```{r}
dt_ssd %>% write_csv(paste0("dt_layton_",Sys.Date(),".csv"))
```
